---
title: "PreventMED - Power Analysis"
format: gfm
---

```{r}
#| include: false
source("./R/dgr-rct_power-simulation.R")
```

# Design and Data Structure

The following sections provide brief descriptions of the data structures for the
primary and secondary outcomes.

## Primary Outcome (SOTIPS)

The primary outcome will be measured at baseline, one months, six months, and
twelve months.

We argue that a 5 point reduction in SOTIPS score after 12 months as a function
of the medication would be clinically meaningful. This corresponds to an average
reduction of 0.42 points per month.

The power simulations presented here are based on the following model.

```{r}
summary(design_lmm)
```

In this model, "measurement" refers to the amount of time spent (in months) in
the trial. The coefficient of primary interest is the interaction term between
time (measurement) and Degarelix.

The variance components of this model are based on preliminary data from a
different trial with a similar population of participants, adjusted to be
slightly more conservative.

# Power Simulation

## Power to Detect Effects on the SOTIPS

With the assumptions of the model described above, the following simulation
estimates the statistical power for the effect of Degarelix over time.

This simulation assumes *n* = 60 in the TAU arm and *n* = 120 in the treatment
arm.

```{r}
power_simulation
```

If we recruit and retain the full planned sample, we will have statistical power
approaching 100% to detect an effect in a magnitude we regard as clinically
significant.

## Accounting for Data/Power Loss

We can also consider how much power we will have if there is a substantial
attrition rate. The following simulation assumes the loss of 33% of the sample.

```{r}
power_simulation_33
```

If we have a dropout rate of 33% (equally across conditions), we will have
approximately 80% power to detect a clinically significant effect. Lower dropout
rates will yield higher levels of power. As such, attrition rates must be quite
severe before the study becomes underpowered to detect relevant effects.
